---
title: "R Cheatsheet"
author: "Roland Widmer, 17-109-646"
date: "2023-05-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Base R

```{r}
age = c(20, 23, 35) 
name = c("Petra", "Peter", "Peterina") 
female = c(TRUE, FALSE, TRUE) 
df = data.frame(age, name, female)  
```

subset(): Select rows and columns of data frame

```{r}
subset(df, age > 23, select=name:female)
```
transform(): Add or overwrite columns in data frame
```{r}
transform(df, male = !female)
```

aggregate(), tapply(), rowsum(), by(): Grouped calculations

```{r}
aggregate(df$age, by = list(df$female), FUN="mean")

tapply(df$age, df$female, mean)

rowsum(df$age, c(1, 1, 2)) # not the same as rowSums
    # group indicates the group for every row of x

by(df$age, df$female, mean) # wrapper of tapply for df

```

ave(): Grouped transformations

```{r}
ave(df$age, df$female, FUN = mean)
```

rbind(), cbind(): Bind rows/columns of data frame/matrix

```{r}
cbind(rep(1,5), rep(2, 5))
rbind(rep(1,5), rep(2, 5))
```


merge(): Join data frames by key

```{r}
profession <- c("Student", "Bar tender", "Fire fighter")

df_prof <- data.frame(name, profession)

merge(df, df_prof, by.x = "name", by.y = "name")
```


expand.grid(): Cross-join lists/data frames

```{r}
x <- 1:10
y <- 1:10
grid <- expand.grid(x, y) # cartesian product
```


head(), tail(): First/last few elements of object

nrow(), ncol(), dim(): Number of rows/columns of data frame/matrix

order(), rank(): Sort indices, ranks

```{r}
order(df_prof$profession, decreasing = TRUE)

df_prof[order(df_prof$profession, decreasing = TRUE),] # sorted 

v <- c(3, 1, 4, 1, 5, 9, 2, 6, 5, 3, 5)

rank(v) # by default: mean!
order(v)
```


rowSums(), rowMeans(): Row-wise sums/means of data frame/matrix

colSums(), colMeans(): Column-wise sums/means data frame/matrix

cumsum(), cummean(): Cumulative sums and means of vector

```{r}
v <- c(1, 2, 3)
cumsum(v)
cummin(v)
```


reshape(): Transposition/Reshaping of data frame (tricky interface) ?

lapply(): Apply function element-wise, e.g., per column of data frame

```{r}
lapply(grid, quantile)
```


str(): Structure of object, e.g., of a data frame

summary(): Summarizes object, e.g., each column in a data frame

```{r}
summary(grid)
```


mean(), median(), sd(), quantile(), min(), max(): Univariate statistics

```{r}
v <- c(1, 2, 3)
median(v)
sd(v)
quantile(v)
```


table(), prop.table(), addmargins(): Absolute and relative counts

```{r}
library(ggplot2)
t <- table(diamonds$color) # count color occurrences
t
addmargins(t)
t <- prop.table(t) # proportions
t

addmargins(t)
```


cor(), cov(): Bivariate statistics

```{r}
x <- c(1, 2, 3)
y <- c(2, 4, 6)
cor(x, y)
cov(x, y)
```


hist(), barplot(), boxplot(), plot(): Some plot functions
```{r}
y <- rnorm(100)
x <- 1:100

plot(x, y)
hist(y)
boxplot(y)
barplot(1:10)
```
Multivariate plots
```{r}
x <- "color"
boxplot(
  reformulate(x, "price"), # we may use ~ notation too
  data = diamonds
)
```

## Tidyverse
Pipe
```{r}
library(magrittr)
diamonds$color %>% 
  table() %>% 
  prop.table() %>% 
  barplot(col = "chartreuse4")
```

select(), filter(), arrange(), mutate(), summary()
```{r}
library(dplyr)
diamonds %>%
  select(c("color", "price")) %>%  # select columns (or color, price)
  filter(price >= mean(price)) %>% 
  arrange(desc(price)) %>% # multiple cols possible
  mutate(expensive = price > 10000) %>% # create new column
  summary()

diamonds %>% 
  summarize(
    med_carat = median(carat), 
    med_price = median(price)
  )
```

-   `group_by()`, `ungroup()`: Groups the rows by levels of one or more columns. Plays well with other functions like `summarize()`or `mutate()`

```{r}
diamonds %>% 
  select(color, price) %>% 
  group_by(color) %>% 
  mutate(mean = mean(price)) %>% # does not change the dimension of the table
  ungroup() %>% 
  head(2)

```

-   `bind_rows()`, `bind_cols()`: Bind data frames by row/column
```{r}
diamonds %>% select(price) %>% head(2) %>% 
  bind_rows(c(price = 500)) %>% 
  bind_cols(genuine = c(TRUE, TRUE, FALSE))
```

-   `left_join()`, `inner_join()`, `right_join`: Join by key
An inner_join() only keeps observations from x that have a matching key in y.

The most important property of an inner join is that unmatched rows in either input are not included in the result.
```{r}
library(dplyr)
name <- c("Peter", "Petra", "Hanswurst")
age <- c(20, 80, 90)
df_age <- data.frame(name, age)

df_prof %>% 
  inner_join(df_age, by="name") # Peterina, Hanswurst is not part of the result
df_prof %>% left_join(df_age, by="name") # Peterina is not part of the result
# left: keeps everything on left side
# right: keeps everything on right side
```

-   `slice()`: Select rows by position
-   `pivot_wider()`, `pivot_longer()`: Reshape/transpose (in "[tidyr](https://tidyr.tidyverse.org/)", not "dplyr")
```{r}
diamonds %>% 
  select(price, carat) %>% 
  pivot_longer(everything()) %>%  # everything selects all variables
  head()

diamonds %>% 
  select(price, clarity, color) %>% 
  pivot_longer(!price, names_to = "key", values_to = "value") %>% 
  head()
```
## ggplot2

+ operator (developed before the pipe operator!)

Bar plot
```{r}
library(ggplot2)

ggplot(data = diamonds, mapping = aes(x = color)) +
  geom_bar(fill = "black")
```

Histogram
```{r}
ggplot(diamonds, aes(x = price)) +
  geom_histogram(bins = 30) + 
  ggtitle("Histogram")
```
Boxplot
```{r}
hstgr <- ggplot(diamonds, mapping = aes(x = color, y = price)) +
  # ylim(c(0, 10000)) + excludes the larger elements, better:
  coord_cartesian(ylim = c(0, 10000)) + 
  geom_boxplot() +
  theme_gray(base_size = 15) # larger text
hstgr
```
Scatterplot
```{r}
p <- ggplot(diamonds, mapping = aes(x = carat, y = price)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() + # logarithmic scales
  geom_smooth() # "regression"
p
```
Grouped plots:
```{r}
p + facet_wrap(~ color)
```
Faceting + pivot_longer is useful!

Use data in color:
```{r}
p <- ggplot(diamonds, mapping = aes(x = carat, y = price, color = clarity)) +
  geom_point(alpha = 0.2) +
  guides(colour = guide_legend(override.aes = list(alpha = 1))) +
  theme(legend.position = "top")
```
Plotly: Make the plots interactive:
```{r}
library(plotly)
hstgr %>%  ggplotly() # magrittr pipe!
```
## Class
```{r}
class(x ~ y) # get class
# generic plot function finds class of x ~ y and class corresponding method, i.e. graphics:::plot.formula (three : for internal methods), other methods: graphics:::plot.factor
```
Our own class
```{r}
# Function that creates an object of class "student"
student <- function(name) {
  out <- list(
    name = name
  )
  class(out) <- "student"
  out
}
me <- student("Michael Mayer")
print.student <- function(x, ...) {
  cat("Hi, I'm", x$name)
  invisible(x) # do not output x itself
}

print(me)
```

