---
title: "Exercises01"
author: "Roland Widmer"
date: "2023-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1


```{r}
library(tidyverse)
library(insuranceData)
library(plotly)
library(scales)

data(dataCar)
```
    a.  Draw barplots of the discrete variables "numclaims", "agecat" (categorized driver age), and "gender".
    
```{r}

ggplot(data = dataCar, mapping = aes(x = numclaims)) +
  geom_bar(fill = "#83b692")

ggplot(data = dataCar, mapping = aes(x = agecat)) +
  geom_bar(fill = "#F9ADA0")

ggplot(data = dataCar, mapping = aes(x = gender)) +
  geom_bar(fill = "#5B3758")
```

    b.  Draw a histogram of the vehicle value "veh_value" (in 10'000 Australian Dollars). Truncate values above 7 (this means: if a value is larger than 7, set it to 7).
    
```{r}
# use pmin instead of rowwise + min (crazy inefficient)
d <- dataCar %>% 
  rowwise %>% 
  mutate(trunc_veh_value = min(veh_value, 7))

p <- ggplot(d, aes(x = trunc_veh_value)) +
  geom_histogram(fill = "#83b692", bins = 30) +
  ggtitle("Histogram of vehicle value") +
  xlab("Price in 10'000 AUD")
p
```
  
    c.  Calculate the average number of claims per level of "agecat" and visualize the result as a scatterplot. Interpret the result.
```{r}
d <- dataCar %>% 
  group_by(agecat) %>% 
  summarise(avg_claims = mean(numclaims))

p <- ggplot(d, mapping = aes(x = agecat, y = avg_claims)) +
  geom_point(color = "black") +
  xlab("Age category, 1=youngest") +
  ggtitle("Scatterplot")
p
```
Interpretation: The higher the age category, the lower the average number of claims. Especially, age categories 5 and 6 have a lower average than the other age categories and age category has a higher average than the others.

    d.  Bin "veh_value" into quartiles and analyze its association with the number of claims as in 1c.
```{r}
d <- dataCar %>%
  mutate(quartile_rank = ntile(veh_value,4)) %>% 
  group_by(quartile_rank) %>% 
  summarise(avg_claims = mean(numclaims))

p <- ggplot(d, mapping = aes(x = quartile_rank, y = avg_claims)) +
  geom_point(color = "black") +
  xlab("Quartile rank, 1=0.25 quartile") +
  ggtitle("Scatterplot")
p

```
Interpretation: The higher the value of the vehicle, the higher the average number of claims.

    e.  Use the "plotly" package to turn the plot from d. interactive.
```{r}
p %>% 
  ggplotly()
```

## 4. Sieve of Eratosthenes
The sieve of Eratosthenes is an ancient algorithm to get all prime numbers up to any given limit $n$, see [Wikipedia](https://en.wikipedia.org/wiki/Sieve_of_Eratosthenes). Write a function `sieve_of_eratosthenes(n)` that returns all prime numbers up to $n$. Benchmark the results for $n = 10^5$ with the package "microbenchmark". Mind your coding style!

Naive approach:
```{r}
sieve_of_eratosthenes <- function(n) {
  df <- data.frame(seq(2,n))
  colnames(df) <- c("x")
  
  p <- 2
  i <- 1
  done <- FALSE
  while(!done) {
    df <- df %>% filter(x %% p != 0 | x == p)
    
    i <- i + 1
    if (i > nrow(df)) {
      done = TRUE
    } else {
      p <- df$x[i]
    }
  }
  df
}
head(sieve_of_eratosthenes(1000))
```
```{r}
sieve_of_eratosthenes_opt <- function(n) {
  v <- seq(2,n)
  
  p <- 2
  i <- 1
  done <- FALSE
  while(!done) {
    v <- v[v %% p != 0 | v == p]
    
    i <- i + 1
    if (i > length(v)) {
      done = TRUE
    } else {
      p <- v[i]
    }
  }
  v
}

# we need to loop only up to sqrt(n)
# debugonce() to go in debug mode
# see solution for C++ implementation with rcpp
# all.equal() to check if two vectors are the same

head(sieve_of_eratosthenes_opt(1000))
```


Comparison
```{r}
library(microbenchmark)
n = 10000
microbenchmark(
  sieve_of_eratosthenes(n),
  sieve_of_eratosthenes_opt(n),
  times = 2
)

```

## 4
In Exercise 1c, we have calculated and plotted the average number of claims per level of "agecat" in the `dataCar` data.

    a.  Write a function `avg_claim_counts(v)` that provides such a visualization for any discrete variable `v`.
  
Source: https://shixiangwang.github.io/tidyeval-chinese/dplyr.html
    
```{r}
avg_claim_counts <- function(v) {
  v <- enquo(v)
  
  d <- dataCar %>% 
  group_by(!!v) %>% 
  summarise(avg_claims = mean(numclaims))

  p <- ggplot(d, mapping = aes(x = !!v, y = avg_claims)) +
    geom_point(color = "black") +
    xlab(v) +
    ggtitle("Scatterplot")
  p
}

avg_claim_counts(agecat)
avg_claim_counts(gender)
avg_claim_counts(area)
```

    b.  Extend this function with a second argument `interactive` to control whether the resulting plot is interactive or not.
    
```{r}
library(plotly)

avg_claim_counts_2 <- function(v, interactive = FALSE) {
  v <- enquo(v)
  
  d <- dataCar %>% 
  group_by(!!v) %>% 
  summarise(avg_claims = mean(numclaims))

  p <- ggplot(d, mapping = aes(x = !!v, y = avg_claims)) +
    geom_point(color = "black") +
    ggtitle("Scatterplot")
  if (interactive) {
    p <- p %>% 
      ggplotly()
  }
  p
}

avg_claim_counts_2(agecat, TRUE)
avg_claim_counts_2(gender, TRUE)
avg_claim_counts_2(area, TRUE)
```
## 5

Extend the "student" class from Section "plot, print, summary" by the optional information "semester". It represents the number of semesters the student is already registered. Add a `summary()` method that would neatly print the name and the semester of the student.

```{r}
# Function that creates an object of class "student"
student <- function(given_name, family_name, semester = NULL) {
  out <- list(
    given_name = given_name,
    family_name = family_name,
    semester = semester
  )
  class(out) <- "student"
  out
}

me <- student("Roland", "Widmer")
me2 <- student("Roland", "Widmer", 11)

summary.student <- function(x, ...) {
  cat(x$given_name, x$family_name, ifelse(x$semester, paste(", Semester:", x$semester), ""))
  invisible(x) # do not output x itself
}

summary(me)
summary(me2)
```

